```{r}
#import packages
library(ape)
library(phangorn)
library(MASS)
library(smacof)
```

```{r}
#simulate data & create distant matrix
set.seed(1)
nsp <- 5
l = 500

#This is the true tree
tree <- rtree(nsp)

#Let's now simulate the sequence alignment
tree <- rtree(5)
write.tree(tree, "tree0.nwk")
system("iqtree2 --alisim alignment0 -m JC -t tree0.nwk -seed 123")
data <- as.phyDat(read.dna("alignment0.phy"))

#We need to estimate the distance matrix for the estimated alignment
dm <- dist.ml(data)
dm[is.na(dm)] <- 0

#Based on the estimated distance matrix, we can infer a quick phylogeny
treeUPGMA <- upgma(dm)
treeNJ <- NJ(dm)
```

Now, are the estimated and true trees similar? We'll use the UPGMA for this.

```{r}
comparePhylo(treeUPGMA, tree, plot = TRUE, force.rooted = FALSE)
```

Let's now perform MDS on the distance matrix

```{r}
#performs multidimensional scaling (MDS) & plot results 
fit_reg <- mds(dm, ndim = 2)
#plot(fit_reg, main = "MDS")
mds_tree <- as.phylo(hclust(dm))
plot(mds_tree)
```

Now, we can compare the MDS-based to the UPGMA and original trees

```{r}
treeUPGMA <- unroot(treeUPGMA)
mds_tree <- unroot(mds_tree)
tree <- unroot(tree)

comparePhylo(treeUPGMA, mds_tree, plot = TRUE, force.rooted = FALSE, use.edge.length = TRUE)
comparePhylo(tree, mds_tree, plot = TRUE, force.rooted = FALSE, use.edge.length = TRUE)
```

Let's now fit an inverse MDS for the MDS tree

```{r}
#performs inverse MDS
fit_in <- inverseMDS(fit_reg$conf)
```

Build the corresponding tree from each of the new matrices generated using `inverseMDS`.

```{r}
fit_in_trees <- lapply(fit_in, function(x) as.phylo(hclust(x)))
```

Now, we need to compare the topology between the original tree and the newly generated ones.

```{r}
x=3
comparePhylo(fit_in_trees[[x]], tree, plot = TRUE, force.rooted = FALSE, use.edge.length = TRUE)
```


Now, we are going to do a one-hot-encoding of the sequence alignment

```{r}
dna <- apply(as.character(data), 1, function(x) paste0(x, collapse = ""))
dna <- toupper(dna)
ohe <- matrix(t(diag(4)[match(unlist(lapply(dna, utf8ToInt)), utf8ToInt("ACGT")), ]), nrow = length(dna), byrow = TRUE)
ohe <- data.frame(ohe)
```

We will drop columns that are rather constant
```{r}
cst <- sapply(ohe, function(x) length(which( x == 1))  )
ohe2 <- ohe[,-which(cst <2)]
```

Now, we will predict the MDS-based vector from the one-hot encoded regression model. Let's do the train/test split first.

```{r}
library(caret)
library(rsample)
library(ggplot2)
fullds <- cbind.data.frame(mds_vector = fit_reg$conf, ohe2)
base_split <- initial_split(fullds, prop = .7)
train <- training(base_split)
test  <- testing(base_split)
```

We will now fit to RF models, each predicting a different dimension of the MDS.

```{r}
library(randomForest)
set.seed(123)

m1 <- randomForest(
  formula = mds_vector.D1 ~ .,
  data    = train
)

m2 <- randomForest(
  formula = mds_vector.D2 ~ .,
  data    = train
)

```

```{r}
iqtree2 --alisim dna -m JC -t tree.nwk -seed 123
```



