```{r}
#import packages
library(ape)
library(phangorn)
library(MASS)
library(smacof)
```

```{r}
#simulate data & create distant matrix
set.seed(1)
nsp <- 5
l = 10000

#This is the true tree
tree <- rtree(nsp)

#Let's now simulate the sequence alignment
data <- simSeq(tree, l = l, type = "DNA", bf = c(.1,.2,.3,.4), Q = 1:6, ancestral = FALSE)

#We need to estimate the distance matrix for the estimated alignment
dm <- dist.ml(data)
dm[is.na(dm)] <- 0

dna <- apply(as.character(data), 1, function(x) paste0(x, collapse = ""))
dna <- toupper(dna)
ohe <- matrix(t(diag(4)[match(unlist(lapply(dna, utf8ToInt)), utf8ToInt("ACGT")), ]), nrow = length(dna), byrow = TRUE)

#Based on the estimated distance matrix, we can infer a quick phylogeny
treeUPGMA <- upgma(dm)
treeNJ <- NJ(dm)
```

Now, are the estimated and true trees similar? We'll use the UPGMA for this.

```{r}
comparePhylo(treeUPGMA, tree, plot = TRUE, force.rooted = FALSE)
```

Let's now perform MDS on the distance matrix

```{r}
#performs multidimensional scaling (MDS) & plot results 
fit_reg <- mds(dm)
plot(fit_reg, main = "MDS")
mds_tree <- as.phylo(hclust(dm))
plot(mds_tree)

fullds <- cbind.data.frame(mds_vector = fit_reg$conf[,1], ohe[,1:5])

summary(lm(mds_vector ~ ., fullds))
```

Now, we can compare the MDS-based to the UPGMA and original trees

```{r}
treeUPGMA <- unroot(treeUPGMA)
mds_tree <- unroot(mds_tree)
tree <- unroot(tree)

comparePhylo(treeUPGMA, mds_tree, plot = TRUE, force.rooted = FALSE, use.edge.length = TRUE)
comparePhylo(tree, mds_tree, plot = TRUE, force.rooted = FALSE, use.edge.length = TRUE)
```

Let's now fit an inverse MDS for the MDS tree

```{r}
#performs inverse MDS
fit_in <- inverseMDS(fit_reg$conf)
```

Build the corresponding tree from each of the new matrices generated using `inverseMDS`.

Now, we need to compare the topology between the original tree and the newly generated ones.











